---
title: "Assignment2A_CodeBase"
format: html
editor: visual
---

## Assignment 2A

The goal of this assignment is to get us familiar with use of SQL, and linking R to SQL. Preliminary work to this was creating a survey of movies for my friends to fill out, that I could then upload into a SQL database to do further work on. This survey contains ratings from 1-5 for all movies with an additional option to rate the movie as "did not see".

## The Code

For the SQL side of this work, I was able to use the postgres GUI to do all of my work, but I extrapolated the SQL from there. I created 3 different documents to hold the native SQL code that was executed behind the GUI.

This was the query for building the table:

"

-- Table: public.Movies

-- DROP TABLE IF EXISTS public."Movies";

CREATE TABLE IF NOT EXISTS public."Movies" ( "Timestamp" text COLLATE pg_catalog."default", "The Odyssey" text COLLATE pg_catalog."default", "Dune 2" text COLLATE pg_catalog."default", "Minecraft Movie" text COLLATE pg_catalog."default", "Barbie" text COLLATE pg_catalog."default", "Backrooms" text COLLATE pg_catalog."default", "Five Nights at Freddys 2" text COLLATE pg_catalog."default", """""" text COLLATE pg_catalog."default", """""2" text COLLATE pg_catalog."default" )

TABLESPACE pg_default;

ALTER TABLE IF EXISTS public."Movies" OWNER to postgres;

"

This was the Query for importing the data:

"

""\\copy public."Movies"("Timestamp", "The Odyssey", "Dune 2", "Minecraft Movie", "Barbie", "Backrooms", "Five Nights at Freddys 2", """""", """""2") FROM 'C:/Users/Nick/DOWNLO~1/UNTITL~1.CSV/UNTITL\~1.CSV' WITH(FORMAT csv, DELIMITER ',', QUOTE '"', ESCAPE '''');"""

"

This was the Query for exporting the data:

"

SELECT \* FROM public."Movies" LIMIT 100

"

I will then load that data into this file using R:

```{r}
Movies <- read.csv("Movies_raw.csv")
```

As you can see this data is pretty messy, so I'm going to clean it up a little before calling this done. I need to remove the first row and change the columns to not all say "the odyssey". weird anomaly from the google forms.

```{r}
library(tidyverse)
```

```{r}
Movies <- Movies[-1,]

```

```{r}
Movies <- rename(Movies,
                "The Odyssey" = The.Odyssey..The.Odyssey.,
                "Dune 2" = The.Odyssey..Dune.2.,
                "Minecraft Movie" = The.Odyssey..Minecraft.Movie.,
                "Barbie" = The.Odyssey..Barbie.,
                "Backrooms" = The.Odyssey..Backrooms.,
                "Five Nights at Freddys 2" = The.Odyssey..Five.Nights.at.Freddy.s.2.)


```

```{r}
Movies <- Movies |>
  select(Timestamp, `The Odyssey`, `Dune 2`, `Minecraft Movie`, Barbie, Backrooms, `Five Nights at Freddys 2`)
Movies
```

## Conclusion

Overall, This assignment was fairly straightforward mainly due to the use of postgres's user friendly GUI. This allowed me to have a low code/no code approach to importing and exporting data, but for this assignment it is still valuable to know what's happening on the back end to allow that. I did not directly connect to SQL from this R document as you can see above, but I was able to import the .csv, and I included the raw SQL above for how you would do this if you were connected.
